---
title: "Predicting AirBnB Rental Prices"
author: "Group 11: Trevor Isaacson, Jonathan Olavarria, Jasmine DeMeyer"
date: "12/10/2021"
output: beamer_presentation #:
    #theme: "Frankfurt"
    #colortheme: "orchid"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
set.seed(445)
knitr::opts_chunk$set(message = FALSE)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(e1071)
library(rmarkdown)
library(glmnet)
library(knitr)
library(leaps)
library(tree)
library(dplyr)
library(caret)
library(gbm)
library(randomForest)
library(GGally)
library(pls)
library(splines)
library(boot)
```

# Motivation

- You are looking for some additional income and decide renting on AirBnB is the best option
- How much should you rent your extra space for?


# Data

- In general, AirBnB data is very open and be easily accessed
- The original dataset is from a past Kaggle competition 
  - Contained over 74,000 individual listings
- For sake of time and processing power, we took a random sample of 17,500 from those 74,000 listings
- They also provided a testing file
- Since the competition is over, we will compile our final predictions on that file using our best model 

# Data

- Consists of 30 variables
- Variables are about the property, property location, the host and host reviews 
- After cleaning and eliminating variables, our data consisted of 22 variables
- 


# Baseline Regression



# Regression Splines/Generalized Additive Models



# PCR and PLS

- 10 Fold Cross-Validation was performed for number of components ranging from 1 to 20.
- The Cross-Validation MSE was used to pick optimal number of components for both models.

# PCR

```{r, echo = FALSE}
#Fit PCR
m0 <- pcr(price~., data = training,ncomp= 20, validation = "CV")
mse <- MSEP(m0)
data.frame(M = mse$comps, mse = t(as.data.frame(mse$val))[, "CV"]) %>%
ggplot() +
  geom_line(aes(M, mse)) +
  geom_point(aes(M, mse)) + ggtitle("Principal Component Regression MSE") +xlab("Number of Components")
```

# PCR Predictions

```{r, include = FALSE}

```


# PLS

```{r,echo = FALSE}
m1 <- plsr(price~., data=training, validation="CV",ncomp=20)

# Plot MSE
mse1 <- MSEP(m1)
data.frame(M = mse1$comps, mse = t(as.data.frame(mse1$val))[, "CV"]) %>%
ggplot() +
  geom_line(aes(M, mse)) +
  geom_point(aes(M, mse))

```



# PLS Predictions

```{r, include = FALSE}

```



# Questions?




## References

